---
title: "p8105_hw2_fw2334"
author: "Frances Williams"
date: "9/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# **PROBLEM 1**

## Creating datasets
#### *Tidying Mr. Trash Wheel dataset*
```{r}
trash_data = read_excel(
  path = "data/TrashWheel2019.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls)
  )
```


#### *Preparing precipitation datasets*
```{r}
precip_2017 = read_excel(
  path = "data/TrashWheel2019.xlsx",
  sheet = "2017 Precipitation",
  range = "A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2017)

    
precip_2018 = read_excel(
  path = "data/TrashWheel2019.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)
```


#### *Joining precipitation datasets*
```{r}
precip_data = 
  full_join(precip_2017, precip_2018, by = c("year", "month", "total")) %>% 
  mutate(
    month = month.name[month]
  )
```

## Description of the datasets
The `trash_data` dataset has `r nrow(trash_data)` observations and includes data on the types of trash collected by Mr. Trash Wheel dumpsters, such as `plastic bottles` and `cigarette_butts`. The combined `precip_data` dataset has `r nrow(precip_data)` observations, and the key variables `year`, `month`, and `total` precipitation were used to merge the 2017 and 2018 datasets. The total amount of precipitation in 2018 was `r sum(pull(precip_2018, total))` inches, and the median number of sports balls in a dumpster in 2017 was `r median(pull(trash_data, sports_balls))`. 



# **PROBLEM 2**

## Creating datasets

#### *pols dataset*
```{r}
pols_df = read_csv(file = "./data/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = month.abb[as.numeric(month)],
    month = str_to_lower(month),
    president = ifelse(prez_dem == 1, "dem", "gop")) %>% 
  select(-prez_gop, -prez_dem, -day)
```

#### *snp dataset*
```{r}
snp_df = read_csv(file = "./data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  select(-day) %>% 
  mutate(
    month = month.abb[as.numeric(month)],
    month = str_to_lower(month)) %>% 
  arrange(year, month)
```

#### *unemployment dataset*
```{r}
unemployment_df = read_csv(file = "./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment") %>% 
  mutate(year = as.character(year)) %>% 
  arrange(year, month)
```

## Joining datasets
```{r}
pols_snp = 
  left_join(pols_df, snp_df, by = c("year", "month"))

pols_snp_unemp = 
  left_join(pols_snp, unemployment_df, by = c("year", "month"))
```

## Description of datasets

The `pols_df` dataset includes data on the number of politicians by political party (dem/gop) and position (sen/rep/gov/prez) by month and year. The `snp_df` dataset contains the closing values of the S&P stock index by month and year. The `unemployment_df` dataset contains percent unemployment by month and year. The merged dataset has `r nrow(pols_snp_unemp)` observations and `r ncol(pols_snp_unemp)` variables; each row of data is for a single month, and the years of data range from `r min(pull(pols_snp_unemp, year))` to `r max(pull(pols_snp_unemp, year))`. Key variables linking the datasets are `year` and `month`.


# **PROBLEM 3**

```{r}
babynames_df = read_csv(file = "./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic")) %>% 
  distinct()
```


## Table 1: Olivia name rank by year and ethnicity
```{r}
olivia_table = 
  babynames_df %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank) %>% 
  select(ethnicity, `2011`, `2012`, `2013`,`2014`, `2015`, `2016`)

knitr::kable(olivia_table)

```

## Table 2: Male name popularity by year
```{r}
malename_table = 
  babynames_df %>% 
  filter(
    gender == "male",
    rank == 1) %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name) %>% 
  select(ethnicity, `2011`, `2012`, `2013`,`2014`, `2015`, `2016`)

knitr::kable(malename_table)
```


## Scatterplot: Male white non-hispanic baby names by rank and count
```{r}
babynames_scatter =
  babynames_df %>%
  filter(
    ethnicity == "white non hispanic", 
    year_of_birth == 2016,
    gender == "male")

ggplot(babynames_scatter, aes(x = rank, y = count)) +
  geom_point()
  
```

